---
title: "Machine Learning - Assignment"
author: "Naveen Krishnan R"
date: "March 31, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Machine Learning Assignment - 4
==================================

This document describes the analysis and outcome of the analysis done on the Assignment for Machine learning. 

The first part is the various libraries that are used for this analysis

```{r Libraried}
library(caret)
library(randomForest)
library(Hmisc)
```

## Reading file

The next part is for reading the csv files provided and creating a training and test data for analysis

```{r read_files }
training_data <- read.csv("pml-training.csv",na.strings = c("NA", "#DIV/0!", ""))
testing_data  <- read.csv("pml-testing.csv" ,na.strings = c("NA", "#DIV/0!", ""))
inTrain  <- createDataPartition(training_data$classe, p=0.75, list=FALSE)
TrainingSet <- training_data[inTrain,]
TestingSet <- training_data[-inTrain,]
dim(TrainingSet)
dim(TestingSet)
```

## Data Preparation

The next part is to read only the relevent data for analysis. This involves removing the N/A and other irrelevent data

```{r Data Prepeareation }
isnacols <- colSums(is.na(TrainingSet)) == 0
TrainingSet <- TrainingSet[,isnacols]
TestingSet <- TestingSet[,isnacols]

dim(TrainingSet)
dim(TestingSet)
set.seed(5602)
# Remove the first 5 cols as it dont have any relationship with data
TrainingSet <- TrainingSet[,-c(1:5)]
TestingSet <- TestingSet[,-c(1:5)]

```

## Prediction Model 1 Rpart Decission Tree

Validate the prediction with Decission Tree algoritm. This uses the train function in the caret package to perform this analysis

```{r Rpart }
# controlRpart <- trainControl(method="cv", number=3, verboseIter=FALSE)
modelRpart <- train(classe ~ ., data=TrainingSet, method="rpart")
modelRpart$finalModel
# Test this in the test data set

confusionMatrix(TestingSet$classe,predict(modelRpart,TestingSet[,-55]))
```



## Prediction Model 2 Random Forest

Validate the prediction with Random forest algoritm. This uses the train function in the caret package to perform this analysis

```{r RF }


controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modelRF <- train(classe ~ ., data=TrainingSet, method="rf",
                          trControl=controlRF)
modelRF$finalModel
# Test this in the test data set

confusionMatrix(TestingSet$classe,predict(modelRF,TestingSet[,-55]))
```




## Prediction Model 3 LDA 

Validate the prediction with LDA. This uses the train function in the caret package to perform this analysis

```{r LDA }



modelLDA <- train(classe ~ ., data=TrainingSet, method="lda")
modelLDA$finalModel
# Test this in the test data set

confusionMatrix(TestingSet$classe,predict(modelLDA,TestingSet[,-55]))
```

## Final Model 

The final Model will be Random forest as it is giving near perfect results. Using the same for the input test data for results

```{r Final test test }


testing_data <- testing_data[,isnacols]

testing_data <- testing_data[,-c(1:5)]
Predict_data <- predict(modelRF,testing_data[,-55])
```

